Factoring with $n+2$ clean qubits and $n-1$ dirty qubits

### 2017-04-22 paper.tex
Fixing issues found by implementing in ProjectQ - Adding controlled variants of more operations - Wire bundles have thick wires in diagrams - Not showing ancilla counters as 'satisfied' on input side of diagrams - Mentioning bit permutation / multi-not operations. - Merging increment diagrams.

\begin{abstract}
\em
We present size-efficient reversible classical circuits for performing various arithmetic operations aided by ancilla in an unknown state.
We improve the number of qubits needed to factor an $n$-bit number from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $2n+1$ qubits, $n-1$ of which can be an unknown state, without increasing the asymptotic size or depth of the circuit.
\end{abstract}

@@

Because one part of a circuit can borrow dirty bits from another part of the same circuit, circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancilla are too far away to be acquired quickly.
Also, when attempting to reduce the number of bits or qubits required by a circuit, converting clean ancilla to dirty ancilla makes a useful intermediate goal.
In this paper, we reduce the number of qubits required to perform period finding from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits.

@@

All constructions assume a 2s-complement representation of integers.
Multi-qubit arithmetic operations always have the least significant bit towards the top.
For clarity, circuit diagrams will divide operations into separate input and output parts, with inputs shaded gray.

@@

The period-finding circuit in figure \ref{fig:period-finding-solo-phase-qubit} is our high-level starting point.
It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations parametrized by previous measurements, and (d) qubit resets.
The only non-trivial operation is (a), the controlled modular multiplication of an $n$-qubit register (requiring $n-1$ dirty ancilla and 1 clean ancilla).

We can perform modular multiplication with multiply-accumulates and a second clean register as in \cite{beauregard2003}.
However, if we want to allow the register to be dirty, it is convenient to pair each modular multiplication with an inverse multiplication on another register (a ``multiply-both-ways").
This change on its own would trash the second register, since at the end of the circuit the second register will have been multiplied by some unknown constant.
But, since the first register is initialized to store 1 and gets multiplied by an inverse constant, we can untrash the second register by multiplying by whatever is in the first register at the end of the circuit.
Figure \ref{fig:period-finding-solo-phase-qubit-explicit-dirty-register} shows this construction.

Be aware that none of our modular circuit constructions guarantee that out-of-range values (i.e. values not below the modulus $R$) are well behaved.
To ensure the second register's value is in range, we use a clean zero'd ancilla for its most-significant-bit (MSB).

\begin{figure}
  \centering
  \makebox[\linewidth]{
    \includegraphics[width=\linewidth]{assets/shor-period-finding.png}
  }
  \caption{\em
	High-level period finding circuit \cite{Shor1999}.
	$R$ is the modulus being queried, $B$ is a randomly chosen base, $n$ is the number of bits needed to store $R$, and $p \in \Theta(N)$ controls the precision of the phase estimation step.
    The triangles indicate how many ancilla are needed``behind the scenes", by our constructions, to perform an operation.
	Recovering the period requires classically processing the sampled output with a continued fractions algorithm.
  }
  \label{fig:period-finding}
\end{figure}

@@

\subsection{Controlled Modular Multiply-Both-Ways}

As in \cite{beauregard2003}, we perform controlled modular multiplication with modular multiply-accumulates into a second register.
However, by using a third multiply-accumulate, we end up with a compound multiply-and-inverse-multiply operation instead of requiring a clean ancilla register.

As with all other modular arithmetic constructions in this paper, we require that the input registers be in range (i.e. have values below $R$).

Suppose that the two registers start in the state $(a, b)$, working modulo $R$.
By multiply-adding the first register times $K$ into the second register, the system advances to $(a, b+aK)$.
Then a multiply-subtract times $K^{-1}$ out of the first register puts the system into the state $(a-bK^{-1}-aKK^{-1}, b+aK)$, which is just $(bK^{-1}, b+aK)$.
Next, we cancel the $b$ term in the second register by multiply-adding the first register times $K$ into it again, leaving $(bK^{-1}, b+aK-bK^{-1} \cdot K)$ which is simply $(bK^{-1}, aK)$.
Finally, we swap to $(aK, bK^{-1})$.
The overall effect is to multiply the first register by $K$ and the second register by $K^{-1}$, as desired.
See figure \ref{fig:controlled-modular-multiply} for the circuit diagram.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-multiply.png}
  \caption{\em
    Performing a controlled modular multiply-both-ways with three controlled modular multiply-accumulates.
    $K$ must have a multiplicative inverse modulo $R$.
  }
  \label{fig:controlled-modular-multiply}
\end{figure}

\subsection{Controlled Modular Multiply-Accumulate}

To perform multiply-accumulation, we use a shift-and-add approach similar to \cite{beauregard2003}.
We right-shift (i.e. divide by $2 {\pmod R}$) the target register $n-1$ times, then begin iteratively left-shifting (i.e. multiplying it by 2) and adding $K$ into the target conditioned on the most significant bit of the input, then the next most significant bit, and so forth.
(The more significant bits go first because their effects must be hit by more left-shifts.)
See figure \ref{fig:controlled-modular-multiply-accumulate}.

A single dirty bit is required by the conditional offset operations, and two dirty bits are required by the modular doublings, but there are more than enough unused bits available to borrow, so the overall circuit doesn't require any dirty bits.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-multiply-accumulate.png}
  \caption{\em
    Reducing a controlled modular multiply-accumulate into the modular equivalent of shift-and-add.
    Assumes $R$ is odd.
  }
  \label{fig:controlled-modular-multiply-accumulate}
\end{figure}

\subsection{Controlled Pivot-Flips}

We will implement both modular doubling and modular offset/addition in terms of ``pivot-flips".
A pivot-flip is an operation that reverses the order of states less than a given pivot value, without affecting other states.
For example, a pivot-flip with the pivot equal to 4 would swap $|0\rangle$ and $|3\rangle$, swap $|1\rangle$ and $|2\rangle$, and leave all other states untouched.

@@

To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - A)$ nearly does what is required: it flips the range below $A$ but unfortunately also flips the range above-and-including $A$.
However, because $x \rightarrow \lnot(x - A)$ is its own inverse and doesn't change whether a given value is below the pivot, the operation can be toggle-controlled by the comparison against $A$.
See figure \ref{fig:controlled-pivot-flip} for the construction.
For states that are less than the pivot $A$, the comparison is true, the ancilla is toggled both times, and exactly one of the controlled subtract-and-inverts will fire.
For other states, the ancilla does not change and so the subtract-and-inverse happens either no times or two times (undoing itself) for a net effect of no effect.

Note that, depending on whether or not the pivot is a compile-time constant or another qubit register, different subtraction constructions are used when reducing further.
This distinction matters because the constant-offset case uses $O(n \lg n)$ gates \cite{haner2016} instead of the $\Theta(n)$ gates used by enregistered addition \cite{takahashi2005}.
Also, the constant-offset case requires an extra dirty bit unless quantum operations are used (e.g. by using the technique demonstrated in figure \ref{fig:bootstrap-ancilla}).

@@

To multiply a register by 2 modulo an odd $R$, we use pivot-flips and subtraction to align the two halves of the input space on the MSB boundaries.
Then we left-rotate.
Note that this trashes out-of-range values larger than $R$ instead of leaving them alone.
See figure \ref{fig:modular-double}.

To invert the doubling, i.e. to divide by 2, recurse and reverse: invert each operation in the construction, then apply the operations in reverse order. 
Recurse-and-reverse works on all circuits presented in this paper (e.g. to perform subtraction instead of addition, or to decrement instead of increment), except period-finding, because the circuits don't generate garbage or use measurement.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/modular-double.png}
  \caption{\em
    Modular doubling.
    Assumes $R$ is odd.
    Requires 2 dirty ancilla, or 1 if using quantum operations.
  }
  \label{fig:modular-double}
\end{figure}

\subsection{Controlled Modular Addition and Offset}

As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition is three pivot flips.
To add $A$ into a register modulo $R$, perform pivot-flips with the pivot at $R-A$, then $R$, then $A$.
See figure \ref{fig:modular-add} for the circuit.

Although our modular addition circuit requires a dirty bit, {\em controlled} modular addition does {\em not}.
See figure \ref{fig:controlled-modular-add}.
However, when performing modular offset (i.e. adding a compile-time constant to a register), we instead require two dirty bits.
See figures \ref{fig:modular-offset} and \ref{fig:controlled-modular-offset}.

Be aware that pivot-flip-based-addition assumes that {\em both} the input and target register are in range.
If the input is greater than or equal to $R$, the pivot-flips will trash the target register instead of adding the input mod $R$ into it.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/mod-add-from-pivot-flip-bars.png}
  \caption{\em
     Modular addition of $A \pmod{R}$ with three pivot flips.
     One at $R-A$, then one at $R$, then one at $A$.
     Requires $A \leq R$ (or else the operation will trash the target).
   }
  \label{fig:mod-add-from-pivot-flip-bars}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/modular-addition.png}
  \caption{\em Non-constant modular addition from pivot-flips.}
  \label{fig:modular-add}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-addition.png}
  \caption{\em
    Controlled modular addition.
    The pivot-flips don't need to be controlled so, unlike the uncontrolled modular addition construction, this construction requires no dirty bits.
  }
  \label{fig:controlled-modular-add}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/modular-offset.png}
  \caption{\em Modular offset.}
  \label{fig:modular-offset}
\end{figure}

@@

\subsection{Addition and Offset}

@@

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/increment-odd.png}
  \caption{\em Odd-sized $O(N)$ increment with 1 dirty ancilla.}
  \label{fig:increment-odd}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/increment-even.png}
  \caption{\em Even-sized $O(N)$ increment with 1 dirty ancilla.}
  \label{fig:increment-even}
\end{figure}

@@

@@

Our main improvements over previous constructions are 1) the use of pivot-flips for modular addition, 2) the use of multiply-both-ways for modular multiplication, and 3) the $O(n)$ incrementer requiring only a single dirty ancilla.

Previous modular addition constructions worked by temporarily storing an is-wraparound-needed comparison in a clean ancilla \cite{takahashi2006, haner2016}.
Pivot flips also require an ancilla (two, actually), but the ancilla can be dirty and in the context of period-finding there are many qubits available to borrow whenever a pivot-flip is needed.
This improvement saves a qubit, reducing the total number of qubits required for period finding to $2n+1$.

Previous modular multiplication constructions used a clean ancilla register \cite{haner2016}.
For these constructions, any garbage in the ancilla register would infect the target register and trash it.
We use an extra multiply-accumulate to cancel out garbage entering the target register, but the ancilla register is still multiplified by the inverse of whatever the target register is multiplied by, creating a "multiply-both-ways" circuit.
We don't want to multiply the ancilla register, since it must be returned to its starting state before our circuit ends, but we can fix the problem with a final cleanup multiplication informed by a measurement of the main register.
This improvement allows $n-1$ of the qubits in the ancilla register, all except the MSB, to be dirty.
The MSB must be zero to ensure values are not larger than the modulus.

@@

We presented Toffoli-based arithmetic circuits that reduce the number of qubits required to perform Shor's algorithm by 1, and the number of {\em clean} qubits required by an additional $n-1$, compared to the best previous constructions.

>>>>

\begin{abstract}
\em
We present asymptotically size-efficient reversible classical circuits for performing various arithmetic operations aided by ancilla in an unknown state.
We improve the number of qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $2n+1$ qubits, $n-1$ of which can be an unknown state, without increasing the asymptotic size or depth of the circuit.
\end{abstract}

@@

Because one part of a circuit can borrow dirty bits from another part of the same circuit, circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancilla are too far away to be acquired quickly.
When attempting to reduce the number of bits or qubits required by a circuit, converting clean ancilla to dirty ancilla makes a useful intermediate goal.
In this paper, at great cost to the constant factors in the gate count and depth, we reduce the number of qubits required to perform period finding from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits without increasing the asymptotic costs.

@@

All constructions assume a 2s-complement representation of integers.
Multi-qubit arithmetic operations always have the least significant bit towards the top.
For clarity, circuit diagrams will divide operations into separate input and output parts, with inputs shaded gray, when applicable.
For completeness, even when the period finding reduction doesn't produce a controlled version of an operation, we provide controllable constructions of that operation that scale well with the number of controls.

@@

The period-finding circuit in figure \ref{fig:period-finding-solo-phase-qubit} is our high-level starting point.
It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations parametrized by previous measurements, and (d) qubit resets.
The only non-trivial operation is (a), the controlled modular multiplication of an $n$-qubit register (requiring $n-1$ dirty ancilla and 1 clean ancilla using our constructions).

We perform modular multiplication with multiply-accumulates and a second clean register as in \cite{beauregard2003}.
However, to allow the second register to be dirty, it is convenient to pair each modular multiplication with an inverse and negated multiplication on another register.
We will refer to this combined operation as a ``bimultiply".

Note that we must preserve the state of the second register, but bimultiplication trashes that state with many multiplications by unknown constants.
However, because the first register is initialized to store 1 and gets multiplied by constants inverse to the ones trashing the second register, we can fix the second register by measuring the first register and using its value to perform a clean-up bimultiplication after the usual end of the circuit.
Figure \ref{fig:period-finding-solo-phase-qubit-explicit-dirty-register} shows this construction.

If the second register contained a value larger than $R$, it would end up trashed despite the final clean-up operation.
Our modular circuit constructions do not guarantee that out-of-range values (i.e. values not below the modulus $R$) are well behaved.
To avoid this problem, we require that the registers have minimal size $n = \lceil \lg_2(R-1) \rceil$ and that the second register's most-significant-bit (MSB) be a clean qubit initialized to zero.

\begin{figure}
  \centering
  \makebox[\linewidth]{
    \includegraphics[width=\linewidth]{assets/shor-period-finding.png}
  }
  \caption{\em
	High-level period finding circuit \cite{Shor1999}.
	$R$ is the modulus being queried, $B$ is a randomly chosen base, $n$ is the number of bits needed to store $R$, and $p \in \Theta(n)$ controls the precision of the phase estimation step.
    The triangles indicate how many ancilla are needed ``behind the scenes", by our constructions, to perform an operation.
	Recovering the period requires classically processing the sampled output with a continued fractions algorithm.
  }
  \label{fig:period-finding}
\end{figure}

@@

\subsection{Modular Bimultiplication}

As in \cite{beauregard2003}, we perform controlled modular multiplication with a second register and modular scaled-multiply-accumulates operations.
However, because the second register is dirty, we need three scaled-multiply-accumulates instead of the usual two.

As with all other modular arithmetic circuits in this paper, we make no guarantees about the circuit's behavior when either of the input registers is out of range (i.e. have values not below $R$).

Suppose that the two registers start in the state $(a, b)$, working modulo $R$.
By multiply-adding the first register times $K$ into the second register, the system advances to $(a, b+aK)$.
Then a multiply-subtract times $K^{-1}$ out of the first register puts the system into the state $(a-bK^{-1}-aKK^{-1}, b+aK)$, which is just $(-bK^{-1}, b+aK)$.
Next, we cancel the $b$ term in the second register by multiply-adding the first register times $K$ into it again, leaving $(-bK^{-1}, b+aK-bK^{-1} \cdot K)$ which is simply $(-bK^{-1}, aK)$.
Finally, we swap to $(aK, bK^{-1})$.
The overall effect is to multiply the first register by $K$ and the second register by $-K^{-1}$, as desired.
See figure \ref{fig:controlled-modular-multiply} for the circuit diagram.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-multiply.png}
  \caption{\em
    Performing a modular bimultiplication with three scaled-multiply-accumulates and a swap.
    $K$ must have a multiplicative inverse modulo $R$.
  }
  \label{fig:controlled-modular-multiply}
\end{figure}

\subsection{Modular Scaled-Multiply-Accumulation}

To perform scaled-multiply-accumulation, we use a shift-and-add approach similar to \cite{beauregard2003}.
We right-shift (i.e. divide by $2 {\pmod R}$) the target register $n-1$ times, then begin iteratively left-shifting (i.e. multiplying it by 2) and adding $K$ into the target conditioned on the most significant bit of the input, then the next most significant bit, and so forth.
(The more significant bits go first because their effects must be hit by more left-shifts.)
See figure \ref{fig:controlled-modular-multiply-accumulate}.

The conditional offset operations need one dirty bit, and the modular doubling need two, but there are more than enough unused bits available to borrow, so the circuit as a whole doesn't require any dirty bits.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-multiply-accumulate.png}
  \caption{\em
    Reducing a modular multiply-accumulate into the modular equivalent of shift-and-add.
    Assumes $R$ is odd.
  }
  \label{fig:controlled-modular-multiply-accumulate}
\end{figure}

To multiply a register by 2 modulo an odd $R$, we use pivot-flips (see section \ref{sec:pivot-flips}) and subtraction to align the two halves of the input space on the MSB boundaries.
Then we finish the job with a left-rotate.
See figure \ref{fig:modular-double}.

To invert the doubling, i.e. to divide by 2 instead of multiplying by 2, recurse and reverse: invert each operation in the construction, then apply the operations in reverse order. 
Except for period finding, all circuits presented in this paper can be inverted (e.g. from addition to subtraction) with recurse-and-reverse.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-double.png}
  \caption{\em
    Controlled modular doubling in $O(c + n \lg n)$ gates, $O(c + n)$ depth, with 2 dirty ancilla.
    $R$ must be odd.
  }
  \label{fig:modular-double}
\end{figure}

\subsection{Modular Addition / Offset}

As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition is three pivot-flips (see section \ref{sec:pivot-flips}).
To add $A$ into a register modulo $R$, perform pivot-flips with the pivot at $R-A$, then $R$, then $A$.
See figure \ref{fig:controlled-modular-add} for the circuit.

Interestingly, modular addition only requires a dirty bit if it's {\em not} controlled.
And when performing modular offset (i.e. adding a compile-time constant to a register), {\em two} dirty bits are required whether or not the operation is controlled.
See figure \ref{fig:controlled-modular-offset}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/mod-add-from-pivot-flip-bars.png}
  \caption{\em
     Modular addition of $A \pmod{R}$ can be done with three pivot flips.
     One at $R-A$, then one at $R$, then one at $A$.
     Requires $A \leq R$ (or else the operation will trash the target).
   }
  \label{fig:mod-add-from-pivot-flip-bars}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-addition.png}
  \caption{\em
    Controlled modular addition.
    The circuit as a whole only needs a dirty bit if no control qubits are available to be borrowed by the uncontrolled pivot-flip operations.
  }
  \label{fig:controlled-modular-add}
\end{figure}

\subsection{Pivot-Flips} \label{sec:pivot-flips}

We implemented both modular doubling and modular offset/addition in terms of a non-standard operation we call ``pivot-flips".
A pivot-flip is an operation that reverses the order of states less than a given pivot value, without affecting other states.
For example, a pivot-flip with the pivot equal to 4 would swap $|0\rangle$ and $|3\rangle$, swap $|1\rangle$ and $|2\rangle$, and leave all other states untouched.

@@

To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - A)$ nearly does what is required: it flips the range below $A$ but unfortunately also flips the range above-and-including $A$.
However, because $x \rightarrow \lnot(x - A)$ is its own inverse, the operation can be ``toggle-controlled'': unless a controlling qubit is toggled between two applications of the operation, it will have no effect or undo itself.
Also, because the operation doesn't change whether a given value is below the pivot, the controlling qubit can be toggled by a comparison of the {\em target} against the pivot $A$.
See figure \ref{fig:controlled-pivot-flip} for the construction.
For states that are less than the pivot $A$, the comparison is true, the ancilla is toggled both times, and exactly one of the controlled subtract-and-inverts will fire.
For other states, the ancilla does not change and so the subtract-and-inverse happens either no times or two times (undoing itself) for a net effect of no effect.

Note that, depending on whether or not the pivot is a compile-time constant or another qubit register, different subtraction constructions are used when reducing further.
This distinction matters because the constant-offset case uses $O(n \lg n)$ gates \cite{haner2016} instead of the $\Theta(n)$ gates used by enregistered addition \cite{takahashi2005}.
Also, the constant-offset case requires an extra dirty bit (unless quantum operations are used, e.g. see figure \ref{fig:bootstrap-ancilla}).

@@

@@

\subsection{Addition / Offset}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-addition.png}
  \caption{\em
  	Reducing controlled addition to uncontrolled addition.
  	Uses $O(N + C)$ gates and depth.
  }
  \label{fig:controlled-addition}
\end{figure}

@@

@@

@@

\subsection{Bit Swaps, Rotations, and Reversals}

Bit permuting operations are easy to overlook in circuits, because they can be emulated by re-labelling qubits.
But some of our circuit diagrams have used controlled bit permutations, requiring actual operations, so we provide the constructions for completeness.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-bit-rotate.png}
  \caption{\em
    A controlled bit rotation / bit swap is three $O(N+C)$ controlled bit-reverses.
  }
  \label{fig:dependencies}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-reverse.png}
  \caption{\em
    An inline controlled bit order reversal using $O(N + C)$ gates.
    Each XOR operation is a series of CNOTs (note that the inputs have opposite endian-ness to the outputs).
    For reversals on an even number of bits, swap the center two bits beforehand then use one of them as the center workspace bit from this construction.
  }
  \label{fig:dependencies}
\end{figure}

\subsection{Multi-Nots}

Several of our constructions have used CNOTs with many controls and many targets.
To avoid paying the overhead of $c$ controls for every target, we use one big CNOT to toggle-control all the others. [[[ FIGURE ]]]
We then use a dirty ancilla (i.e. one of the other targets) to efficiently reduce the big CNOT into Toffolis.

@@

We tested individual constructions in Quirk [[[ https://github.com/Strilanc/Quirk ]]] and the full construction down to Toffoli gates in ProjectQ [[[ https://github.com/ProjectQ-Framework/ProjectQ ]]] [[[ link to code ]]].

Our main improvements over previous constructions are 1) the use of pivot-flips for modular addition, 2) the use of bimultiplication for modular multiplication, and 3) the $O(n)$ incrementer requiring only a single dirty ancilla.

Previous modular addition constructions worked by temporarily storing an is-wraparound-needed comparison in a clean ancilla \cite{takahashi2006, haner2016}.
Pivot flips also require an ancilla (two, actually), but the ancilla can be dirty and, in the context of period-finding, there's always qubits available to borrow whenever a pivot-flip is needed.
This improvement saves a qubit, reducing the total number of qubits required for period finding from $2n+2$ to $2n+1$.

Previous modular multiplication constructions required a clean ancilla register \cite{haner2016}, where any garbage in the ancilla register would have infected the target register and trashed it.
We deal with the trash with an extra multiply-accumulate per modular multiplication, and cancel the overall effects on the ancilla register with a clean-up multiplication informed by measuring the main register at the end of the circuit.
This improvement allows $n-1$ of the qubits in the ancilla register  to be dirty (only the MSB must be zero; to ensure the register's value is in range).

@@

We presented Toffoli-based arithmetic circuits that reduce the smallest known number of qubits required to perform Shor's algorithm by 1, and allowed an additional $n-1$ qubits to be dirty.
